---
title: "cciea_data"
author: "Raine Detmer"
date: "2023-12-16"
output: html_document
---

code for making supplemental figure with histograms of CCIEA indicator time series lengths and coefficients of variation


IEA indicators data downloaded on 9/17/2023 from
https://www.integratedecosystemassessment.noaa.gov/regions/california-current/california-current-iea-indicators 


```{r}

# list containing the links to all the data sets

links <- list("https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SIMP.csv?time,simpson_diversity&population=%22Coastwide%22", # coastwide gf S div
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SIMP.csv?time,simpson_diversity&population=%22North%22",  #north gf S div
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SIMP.csv?time,simpson_diversity&population=%22South%22", #south gf S div
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SP_RICH.csv?time,species_richness&population=%22Coastwide%22", #gf sp rich coastwide
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SP_RICH.csv?time,species_richness&population=%22North%22", # gf sp rich north
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SP_RICH.csv?time,species_richness&population=%22South%22", #gf sp rich south
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Adult%20Anchovy%22", # adult anchovy
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Adult%20Sardine%22", # adult sardine
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Market%20squid%22", # market squid
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Jack%20mackerel%22", # jack mackerel
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Juvenile%20chum%22", #juv chum
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Total%20Krill%22" , # krill
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22anchovy%22", # larval anchovy
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22CA%20smoothtongue%22", # larval CA smoothtongue
             "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22croakers%22", # larval croakers
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22eared%20blacksmelt%22", # larval eared blacksmelt
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22English%20sole%22", # larval english sole
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22Jack%20mackerel%22", # larval jack mackerel
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22market%20squid%22", # larval market squid
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22northern%20lampfish%22", # larval northern lampfish
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22hake%22", # larval hake
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22Pacific%20mackerel%22", # larval mackerel
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22rockfishes%22", # laral rockfish
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22sanddabs%22", # larval sanddab
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22sardine%22", # larval sardine
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22slender%20sole%22", # larval slender sole
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Market%20Squid%22", # market squid ccc
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Total%20Myctophids%22", # myctophids
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Northern%20anchovy%22", # N anchovy
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Octopus%22", # octopus,
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Pompano%22", # Pa pompano
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Pacific%20chub%20mackerel%22", # Pa chub mackerel
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Pacific%20hake%22", # Pa hake
             "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Pacific%20herring%22", # Pa herring
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Pacific%20sardine%22", # Pa sardine
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Pacific%20spiny%20dogfish%22", # Pa spiny dogfish
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Pyrosomes%22", # pyrosomes CCC
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Pyrosome%22", # pyrosomes NCC
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Salps%22", # salps
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBS.csv?time,relative_abundance&larval_species_group=%22southern%20mesopelagics%22", # southern mesopelagics
             "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Subyearling%20Chinook%22", #subyearling chinook
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Surf%20smelt%22", # surfsmelt
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22Thetys%22", # thetys
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22Whitebait%20smelt%22", # whitebait smelt
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22YOY%20Anchovy%22", # yoy anch
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22YOY%20Pacific%20Hake%22", # yoy hake CCC
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22YOY%20Pacific%20hake%22", # yoy hake NCC
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22YOY%20Rockfish%22", # yoy rockfish
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22YOY%20Sanddabs%22", # yoy sanddabs
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBC.csv?time,mean_cpue&species_group=%22YOY%20Sardine%22", # yoy sardine
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FPN.csv?time,prevalence&species_group=%22YOY%20Rockfish%20(spp.)%22", # yoy rockfish NCC
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Yearling%20Chinook%22", # yearling chinook
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Yearling%20Coho%22", # yearling coho
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Juvenile%20sockeye%22", #yearling sockeye
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_COP.csv?time,biomass_anomaly&cohort=%22NorthernBiomassAnomaly%22", # N copepod biomass anomaly
             "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_COP.csv?time,biomass_anomaly&cohort=%22SouthernBiomassAnomaly%22", # S copepod biomass anomaly
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_CRAB_FINF.csv?time,biomass_ratio&population=%22Coastwide%22", #fin to crab biomass ratio coastwide
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_CRAB_FINF.csv?time,biomass_ratio&population=%22North%22", # fin to crab north
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_CRAB_FINF.csv?time,biomass_ratio&population=%22South%22", # fin to crab south
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_MTL.csv?time,mean_trophic_level&population=%22Coastwide%22", # groundfish mean trophic level coastwide
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_MTL.csv?time,mean_trophic_level&population=%22North%22", # gf mean trophic level north
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_MTL.csv?time,mean_trophic_level&population=%22South%22", # gf mean trophic level south
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SCAV_RAT.csv?time,biomass_ratio&population=%22Coastwide%22", # scavenger biomass ratio coastwide
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SCAV_RAT.csv?time,biomass_ratio&population=%22North%22", # scavenger biomass ration north
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_SCAV_RAT.csv?time,biomass_ratio&population=%22South%22", # scavenger biomass ratio south
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Aequorea%20Water%20Jelly%22",# Aequorea jelly
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Moon%20jelly%22", # aurelia moon jelly
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Chrysaora%20Sea%20Nettle%22", # sea nettle
              "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_EI_FBN.csv?time,mean_density&species_group=%22Egg%20yolk%20jelly%22" # egg yolk jelly
              ) 


```


```{r}
# names of all of the datasets in the above list

dt_names <- c("coastwide groundfish Simp div" , "north gf S div" , "south gf S div" , "gf sp rich coastwide" , "gf sp rich north" , "gf sp rich south" , "adult anchovy" , "adult sardine" ,"market squid NCC", "jack mackerel" , "juv chum" , "krill" , "larval anchovy" , "larval CA smoothtongue" , "larval croakers",  "larval eared blacksmelt",  "larval english sole",  "larval jack mackerel", "larval market squid", "larval northern lampfish", "larval hake", "larval mackerel", "larval rockfish", "larval sanddab" , "larval sardine", "larval slender sole", "market squid ccc", "myctophids (lanternfish)" , "N anchovy", "octopus", "PA pompano", "PA chub mackerel", "PA hake",  "PA herring", "PA sardine", "PA spiny dogfish" , "pyrosomes CCC" , "pyrosomes NCC" , "salps",  "southern mesopelagics" , "subyearling chinook" , "surfsmelt" , "thetys",  "whitebait smelt", "yoy anch", "yoy hake CCC" , "yoy hake NCC",  "yoy rockfish CCC" , "yoy sanddabs" , "yoy sardine" , "yoy rockfish NCC" , "yearling chinook", "yearling coho", "yearling sockeye" , "N copepod biomass anomaly" ,  "S copepod biomass anomaly" , "crab to fin biomass ratio coastwide", "crab to fin north", "crab to fin south", "groundfish mean trophic level coastwide", "gf mean trophic level north" , "gf mean trophic level south" , "scavenger biomass ratio coastwide" , "scavenger biomass ratio north" , "scavenger biomass ratio south", "Aequorea jelly", "aurelia moon jelly", "sea nettle", "egg yolk jelly") 


```

```{r}
# check lengths are the same
length(links)

length(dt_names)

```


```{r}

# download the data
dt <- vector(mode = "list", length = 1)

for(i in 1:length(links)){
  
  dti <- read.csv(links[[i]])
  dti <- dti[-1, ] # remove header row
  
  dt[[i]] <- dti
  
 # print(colnames(dti))
  
}


```


```{r}

# get summary statistics
ts_lengths <- rep(NA, length(dt)) # time series length
cv_vals <- rep(NA, length(dt)) # coeff of variation
mean_vals <- rep(NA, length(dt)) # mean of response variable
sd_vals <- rep(NA, length(dt)) # sd of response variable
min_vals <- rep(NA, length(dt)) # min value of response variable

for(i in 1:length(dt)){
  dti <- dt[[i]]
  
  #yvals1 <- as.numeric(dti[ ,2])
  # scale to be between 0 and max (so negative values don't mess up cv)
 # yvals <- (yvals1-min(yvals1, na.rm = T))/(max(yvals1, na.rm = T)-min(yvals1, na.rm = T))
  
  yvals <-  as.numeric(dti[ ,2])
  
  ts_lengths[i] <- length(which(is.na(yvals)==F))
  cv_vals[i] <- sd(yvals, na.rm = T)/mean(yvals, na.rm = T)
  mean_vals[i] <- mean(yvals, na.rm = T)
  sd_vals[i] <- sd(yvals, na.rm = T)
  min_vals[i] <- min(yvals, na.rm = T)
}

cciea <- data.frame(
  ts_length = ts_lengths,
  cv = cv_vals,
  mean = mean_vals,
  sd = sd_vals,
  min = min_vals,
  dt_name = dt_names
)

#View(cciea)
#View(dt[[55]]) # copepod biomass anomalies are measured approx monthly so don't count these ts lengths


```



plot histograms of time series lengths (excluding the copepod anomalies which are measured monthly) and coefficients of variations (excluding indicators whose values were both positive and negative)

```{r}

#length(which(cciea$ts_length==25))

pdf("figurepdfs/cciea_sim_pars.pdf", width = 4.5, height = 2)
par(mfrow = c(1, 2))
par(mar=c(2, 1.25, 0.5, 0.75), oma = c(1, 1.2, 1.2, 0.1))
hist(cciea$ts_length[which(cciea$dt_name !="N copepod biomass anomaly" & cciea$dt_name != "S copepod biomass anomaly")], n = 20, xlab = NA,ylab = NA, las = 1, main = NA, xaxs = "i", yaxs = "i",xlim = c(15, 35), ylim = c(0, 40), xaxt = "n", yaxt = "n")
axis(side = 1, at = c(15, 20, 25, 30, 35), padj = -1.75, cex.axis = 0.75, tcl = -0.25)
axis(side = 2, at = c(0, 10, 20, 30, 40), labels = c(NA, 10, 20, 30, 40), las = 1, cex.axis = 0.75, tcl = -0.25, hadj = -0.01)
mtext(side = 1, "Years in time series", line = 1, cex = 0.85)
mtext(side = 2, "Frequency", line = 1.5, cex = 0.85)
mtext(side = 3, "Time series length", adj = 0, line = 0.5, cex = 1)

hist(cciea$cv[which(cciea$min >=0)], n = 20, xlab = NA, ylab = NA, main = NA, las = 1, xaxs = "i", yaxs = "i", xaxt = "n", yaxt = "n", ylim = c(0, 15))
axis(side = 1, at = c(0, 1, 2, 3, 4, 5), padj = -1.75, cex.axis = 0.75, tcl = -0.25)
axis(side = 2, at = c(0, 5, 10, 15), labels = c(NA, 5, 10, 15), las = 1, cex.axis = 0.75, tcl = -0.25, hadj = -0.01)
mtext(side = 1, "Coefficient of variation (cv)", line = 1, cex = 0.85)
#mtext(side = 2, "Frequency", line = 2.5)
mtext(side = 3, "Observation error", adj = 0, line = 0.5, cex = 1)
#mtext(side = 3, "CCIEA indicator time series CVs \n(excluding indicators w/ neg. values)")

dev.off()

```

look more closely at the CV and time series length values

```{r}

# check which indicators had large cv's
cciea[which(cciea$min >=0 & cciea$cv > 1), ]

```


some of these are large bc the mean is small, e.g., for YOY Hake NCC, the mean is 0.0016 and the sd is only 0.008

check the smaller end of the cv values:

```{r}
hist(cciea$cv[which(cciea$min >=0 & cciea$cv < 2)], n = 40, xlab = "cv", main = "CCIEA time series CVs (values <2 only)", las = 1)

hist(cciea$cv[which(cciea$min >=0 & cciea$cv < 0.2)], n = 20, xlab = "cv", main = "CCIEA time series CVs (values <0.2 only)", las = 1)

cciea[which(cciea$min >=0 & cciea$cv < 0.2),] # everything less than 0.2 is a derived value (diversity and trophic structure indicators)

mean(cciea$cv[which(cciea$min >=0 & cciea$cv < 0.2)]) # mean cv of these is 0.05
mean(cciea$cv[which(cciea$min >=0 & cciea$cv > 0.2 & cciea$cv < 1)]) # 0.63
median(cciea$cv[which(cciea$min >=0 & cciea$cv > 0.2 & cciea$cv < 1)]) # 0.65

mean(cciea$cv[which(cciea$min >=0)]) # overall mean is 0.83 

median(cciea$cv[which(cciea$min >=0)]) # over all median is 0.65

quantile(cciea$cv[which(cciea$min >=0)], 0.5)# 0.65



```


the overall median cv is 0.65, and note the smallest values (<0.2; have a mean of 0.05) are all derived values (diversity and trophic structure indicators). 

time series lengths:

```{r}

min(cciea$ts_length[which(cciea$dt_name !="N copepod biomass anomaly" & cciea$dt_name != "S copepod biomass anomaly")]) # 18

max(cciea$ts_length[which(cciea$dt_name !="N copepod biomass anomaly" & cciea$dt_name != "S copepod biomass anomaly")]) # 33

mean(cciea$ts_length[which(cciea$dt_name !="N copepod biomass anomaly" & cciea$dt_name != "S copepod biomass anomaly")]) # 24.5

median(cciea$ts_length[which(cciea$dt_name !="N copepod biomass anomaly" & cciea$dt_name != "S copepod biomass anomaly")]) # 24

```


most of time series are 25 years (low is 18 and max is 33)
